# Introduction

This R Markdown document demonstrates how to create a bar plot to visualize category grouped by MSI_phenotype. Replace the example data with your dataset to create custom plots.

```{r define-colors}
# Define category  colors to be used in the plot
sig_colors <- c("#7AA6DCFF", "#0073C2FF", "#FFA200", "#A52206", "#3B3B3B", "#3B3B3BFF")
```

## Example Dataset

Here is an example dataset that mimics the structure of your data. Ensure your dataset has the following columns:
- `Tumor_Sample_Barcode`: Unique identifier for each sample
- `MSI_phenotype`: Grouping variable (e.g., MSI-H, MSS/TMB-H, MSS/TMB-L)
- `category `: Category. For example (T cell, B cell..)
- `value`: Numeric value for each category (for example amount of cells) 

```{r example-data}
# Create example dataframe
example_data <- data.frame(
  Tumor_Sample_Barcode = rep(paste0("Sample", 1:10), each = 3),
  MSI_phenotype = rep(c("MSI-H", "MSS/TMB-H", "MSS/TMB-L"), times = c(2, 2, 2)),
  category  = rep(c("Sig1", "Sig2", "Sig3"), 10),
  value = runif(30, 0, 1)
)

# Preview the example data
head(example_data)
```

## Data Preparation

```{r data-preparation}
# Prepare data
prepared_data <- example_data %>%
  dplyr::ungroup()

# Convert Tumor_Sample_Barcode to a factor
prepared_data$Tumor_Sample_Barcode <- factor(
  prepared_data$Tumor_Sample_Barcode,
  levels = unique(prepared_data$Tumor_Sample_Barcode)
)
```

## Plotting the Data
```{r create-plot}
# Create the plot
p <- ggplot2::ggplot(prepared_data, ggplot2::aes(x = Tumor_Sample_Barcode, y = value, fill = category )) +
  ggplot2::geom_bar(stat = "identity", position = "stack") +
  ggplot2::theme_minimal() +
  ggplot2::labs(
    y = "Exposure",
    fill = "category ",
    x = ""
  ) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_blank(),  # Remove x-axis text
    axis.ticks.x = ggplot2::element_blank(),  # Remove x-axis ticks
    legend.position = "top",  # Place legend on top
    text = ggplot2::element_text(size = 24),  # Increase all text size
    axis.title = ggplot2::element_text(size = 24),  # Increase axis titles size
    legend.title = ggplot2::element_text(size = 24),  # Increase legend title size
    legend.text = ggplot2::element_text(size = 24),  # Increase legend text size
    strip.text = ggplot2::element_text(size = 24, angle = 0, hjust = 0.5)  # Rotate and size facet labels
  ) +
  ggplot2::scale_fill_manual(values = sig_colors) +
  ggplot2::facet_grid(. ~ MSI_phenotype, scales = "free_x", space = "free_x")  # Free x-axis space based on number of samples

# Display the plot
print(p)
```
