---
title: "Using ggpubr::stat_pvalue_manual()"
output: html_document
date: "2025-01-11"
---

```{r setup, include=FALSE}
library(dplyr) # dplyr is required for %>%
```

## Introduction

This document demonstrates the use of `ggpubr::stat_pvalue_manual()` for adding p-values to boxplots. First, we show how to add p-values to a single boxplot, and then how to use Cell_Types with statistical annotations.

---

## Example Data
```{r create-data}
set.seed(123)
mock_data <- data.frame(
  MSI_phenotype = rep(c("MSS/TMB-H", "MSS/TMB-L", "MSI-H"), each = 40),
  Value = c(
    rnorm(10, mean = 10, sd = 1),
    rnorm(100, mean = 2, sd = 1),
    rnorm(10, mean = 15, sd = 1)  
  )
)
mock_data$Cell_Type <- rep(c("CD8", "CD4"), each = 60)
mock_data
```
Order MSI_phenotype in specific order
```{r}
mock_data <- mock_data %>%
  mutate(MSI_phenotype = factor(
    MSI_phenotype,
    levels = c("MSI-H", "MSS/TMB-H", "MSS/TMB-L"),
    ordered = TRUE
  ))
```

---

## Adding p-values to a single plot

We perform an ANOVA followed by Tukey's HSD test to compute p-values for the differences between MSI_phenotypes. These are then added to a boxplot.
```{r}
cd8_mock_data <- mock_data # Lets say that this data have only 1 cell
```

```{r single-plot}
# Perform ANOVA and Tukey's HSD test
stat.test <- cd8_mock_data %>%
  aov(data = ., Value ~ MSI_phenotype) %>%           # Perform ANOVA to test group differences
  rstatix::tukey_hsd() %>%                           # Perform Tukey's HSD for pairwise comparisons
  rstatix::adjust_pvalue(method = "bonferroni") %>%  # Adjust p-values using the Bonferroni correction method
  rstatix::add_significance("p.adj")                 # Add significance levels based on adjusted p-values

# Add y positions for plotting
stat.test <- as_tibble(stat.test) %>%
  rstatix::add_y_position(                          # Calculate y-axis positions for p-value brackets
    fun = "max",                                    # Place brackets at the maximum value of the groups
    scales = "free",                                # Allow scales to vary if used in a faceted plot
    formula = Value ~ MSI_phenotype,                # Define the relationship being tested
    data = mock_data[, c("Value", "MSI_phenotype")] # Use the corresponding data columns
  )

# Add an empty column for compatibility with ggpubr
# sggpubr requires this column to avoid an error during plotting
stat.test$MSI_phenotype <- ""
```

Add p-value to the plot
```{r}
# Plot
p <-
  ggplot2::ggplot(cd8_mock_data,
                  ggplot2::aes(x = MSI_phenotype, y = Value, fill = MSI_phenotype)) +
  ggplot2::geom_boxplot(alpha = 0.7) +  # Add boxplot
  ggplot2::geom_jitter(
    position = ggplot2::position_jitterdodge(dodge.width = 1),
    size = 2,
    alpha = 0.5,
    shape = 21,
    color = "black"
  ) +  # Add jittered points
  ggplot2::labs(fill = "MSI/TMB", x = "", y = "Count") +
  ggplot2::theme_classic()  +
  ggpubr::stat_pvalue_manual(stat.test, label = "p.adj.signif", tip.length = 0.01, size = 5, bracket.size = 1)

# Display the plot
print(p)
```

## Adding p-values in Cell_Types

Here, we compute separate ANOVA tests for each feature (Cell_Type) and annotate the p-values in each panel.

```{r Cell_Type-plot}
# Perform ANOVA and Tukey's HSD test for each Cell_Type
res.stat <- data.frame()  # Initialize an empty dataframe to store results

# Loop through each unique Cell_Type to calculate p-values
for (group in unique(mock_data$Cell_Type)) {
  # Subset data for the current Cell_Type
  tmp <- mock_data[mock_data$Cell_Type == group,]
  
  # Perform statistical tests
  tmp.stat <- tmp %>%
    aov(data = ., Value ~ MSI_phenotype) %>%            # Perform ANOVA to check for differences between groups
    rstatix::tukey_hsd() %>%                           # Perform Tukey's HSD for pairwise comparisons
    rstatix::adjust_pvalue(method = "bonferroni") %>%  # Adjust p-values using the Bonferroni method
    rstatix::add_significance("p.adj")                # Add significance labels based on adjusted p-values
  
  # Add the current Cell_Type to the result for tracking
  tmp.stat$Cell_Type <- group
  
  # Add xy positions for plotting brackets
  tmp.stat <- as_tibble(tmp.stat) %>%
    rstatix::add_xy_position(
      x = "MSI_phenotype",
      fun = "max",                     # Position brackets at the maximum value in the plot
      scales = "free",                 # Allow each facet to scale independently
      formula = Value ~ MSI_phenotype, # Formula for the relationship being tested
      data = tmp[, c("Value", "MSI_phenotype")]
    )
  
  # Append the results to the main dataframe
  if (nrow(res.stat) == 0) {
    res.stat <- tmp.stat
  } else {
    res.stat <- rbind(res.stat, tmp.stat)
  }
}

# Convert the results to a tibble for use in ggpubr::stat_pvalue_manual
stat.test <- as_tibble(res.stat)
```

```{r}
# Plot with statistical annotations
p <- ggpubr::ggboxplot(
  mock_data,
  x = "MSI_phenotype",
  y = "Value",
  fill = "MSI_phenotype",
  xlab = FALSE,
  ylab = FALSE,
  add = "jitter",  # Add jittered points to show individual data points
  add.params = list(
    shape = 1,  # Circle shape for points
    size = 1,   # Size of points
    alpha = 0.5 # Transparency
  )
) +
  ggplot2::labs(fill = "MSI/TMB", x = "", y = "") +
  ggplot2::facet_wrap(~ Cell_Type, scale = "free") +  # Create separate facets for each Cell_Type
  ggpubr::stat_pvalue_manual(
    stat.test,                     # Use the calculated statistics
    label = "p.adj.signif",        # Label brackets with significance levels (adjusted p-values)
    hide.ns = TRUE,                # Hide non-significant comparisons
    tip.length = 0.01,             # Adjust tip length for brackets
    size = 5,                      # Font size for labels
    bracket.size = 1               # Line thickness for brackets
  ) +
  ggplot2::theme_classic() +
  ggpubr::font("xy.text", size = 15) +
  ggplot2::theme(
    strip.text.x = ggplot2::element_text(size = 20),  # Adjust facet label font size
    axis.text.x = ggplot2::element_blank(),           # Remove x-axis text
    axis.ticks.x = ggplot2::element_blank(),          # Remove x-axis ticks
    legend.position = "top",                          # Place legend on top
    legend.title = ggplot2::element_text(size = 20),  # Adjust legend title font size
    legend.text = ggplot2::element_text(size = 18)    # Adjust legend text font size
  )

# Display the plot
p

```

